Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-28 src/tex/ms.tex
Final typos

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective/loss function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of the optimized waveforms with the original waveforms. We also show how the
optimization result differs as a function of the source parameters. In
Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
parameterization affects the accuracy of the model.

@@

@@

@@

\begin{figure}
	\script{loss.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/loss.pdf}
	\caption{Loss functions against the number of iterations. The vertical
	dotted line indicates minimum of the validation loss, which is where we stop
	the optimization.}
	\label{fig:loss}
\end{figure}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the
	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
	the training loss, green triangles represent extra training data, and the 
	orange dots represent the waveforms used for validation.}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, two SXS waveforms are analogues to two of the BAM NR simulations used in the original calibration (SXS:BBH:1417 as A7 and SXS:BBH:1418 as A9 in \citep{khan2016frequency}). The rest of the SXS waveforms are also used in the original IMRPhenomD calibration.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

\begin{figure*}[t]
	\script{0154.py}
	\centering
	\includegraphics[width=\textwidth]{figures/0154.pdf}
	\caption{Comparison between original and optimized IMRPhenomD waveforms.
	Here we show the SXS:BBH:0154 NR waveform, which has a mass ratio of $q=1$ and spins
	$\chi_1=\chi_2=-0.8$. The original mismatch is $2.8\times10^{-4}$ and
	the optimized mismatch is $5.3\times10^{-5}$. \textit{Top}: Here we show the
	amplitude (left) and phase (right) of the NR, original IMRPhenomD, and optimized IMRPhenomD
	waveform. \textit{Bottom}: Here we show the relative error between the NR and
	IMRPhenomD waveform ampltiudes (left) as well as the absolute error of the phases between the NR and
	IMRPhenomD waveforms (right).}
	\label{fig:0154}
\end{figure*}

@@

\begin{figure}[t]
	\script{q148.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148.pdf}
	\caption{Distributions of waveform mismatches calculated using 
	$\mathcal{L}_{\mathrm{ave}}$. We use the training waveforms listed in 
	Tab.~\ref{tab:q148} and	mismatches are weighted with a constant PSD. 
	The dashed lines represent the median of the distributions, which decreased by 45.3\% during optimization.}
	\label{fig:q148}
\end{figure}
\begin{figure}[t]
	\script{q148_q1248_compare.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
	\caption{Distributions of $\log_{10}$ difference in mismatch. The red
	distribution uses the training waveforms listed in
	Tab.~\ref{tab:q148} while the green distribution uses waveforms
	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
	using a constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
	The dashed lines represent the median of the distributions, which decreased by 10.4\% during optimization.}
	\label{fig:q148_q1248_compare}
\end{figure}

@@

\begin{figure*}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\textwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches for both split region optimization and all region optimization in all 4 
	regions. We use a constant PSD to calculate the mismatch and
	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distributions are indicated by dashed
	lines.}
	\label{fig:all_quadrants}
\end{figure*}

\begin{figure}[t]
	\script{ps_q148_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
	\caption{Combined distribution of both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
	\label{fig:ps_q148_quadrant}
\end{figure}

>>>>

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective/loss function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of the optimized waveforms with the original waveforms. We also show how the
optimization result differs as a function of the source parameters. Finally, in
Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
parameterization affects the accuracy of the model. 
Note that throughout this paper we use the terms recalibration and optimization interchangeably.

@@

\begin{figure}
	\script{loss.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/loss.pdf}
	\caption{Loss functions against the number of iterations. The vertical
	dotted line indicates the minimum of the validation loss, which is where we stop
	the optimization.}
	\label{fig:loss}
\end{figure}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the
	$q-\chi_{\rm PN}$ space. The blue stars represent the waveforms used to compute
	the training loss, green triangles represent extra training data, and the 
	orange dots represent the waveforms used for validation.}
	\label{fig:intrin_space}
\end{figure}

@@

@@

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of NR waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, two SXS waveforms are analogues to two of the BAM NR simulations used in the original calibration (SXS:BBH:1417 as A7 and SXS:BBH:1418 as A9 in \citep{khan2016frequency}). The rest of the SXS waveforms are also used in the original IMRPhenomD calibration.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

\begin{figure*}[t]
	\script{0154.py}
	\centering
	\includegraphics[width=\textwidth]{figures/0154.pdf}
	\caption{Comparison between original and optimized IMRPhenomD waveforms.
	Here we show the SXS:BBH:0154 NR waveform, which has a mass ratio of $q=1$ and spins
	$\chi_1=\chi_2=-0.8$. The original mismatch is $2.8\times10^{-4}$ and
	the optimized mismatch is $5.3\times10^{-5}$. \textit{Top}: Here we show the
	amplitude (left) and phase (right) of the NR, original IMRPhenomD, and optimized IMRPhenomD
	waveforms. \textit{Bottom}: Here we show the relative error between the NR and
	IMRPhenomD waveform ampltiudes (left) as well as the absolute error of the phases between the NR and
	IMRPhenomD waveforms (right).}
	\label{fig:0154}
\end{figure*}

@@

\begin{figure}[t]
	\script{q148.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148.pdf}
	\caption{Distributions of mismatches before and after optimization using the 
	$\mathcal{L}_{\mathrm{ave}}$ loss function. Mismatches are calculated using the training waveforms listed in 
	Tab.~\ref{tab:q148} and	are weighted with a constant PSD. 
	The dashed lines represent the median of the distributions, which decreased by 45.3\% during optimization.}
	\label{fig:q148}
\end{figure}
\begin{figure}[t]
	\script{q148_q1248_compare.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
	\caption{Distributions of $\log_{10}$ difference in mismatch. The red
	distribution uses the training waveforms listed in
	Tab.~\ref{tab:q148} while the green distribution uses waveforms
	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
	using a constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
	The dashed lines represent the median of the distributions, which decreased \te{by an additional 10.4\% during optimization with the additional data.}}
	\label{fig:q148_q1248_compare}
\end{figure}

@@

\begin{figure*}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\textwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches for both split region optimization and all region optimization in all 4 
	regions. We use a constant PSD to calculate the mismatch and
	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distribution are indicated by dashed
	lines.}
	\label{fig:all_quadrants}
\end{figure*}

\begin{figure}[t]
	\script{ps_q148_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
	\caption{Combined mismatch distributions for both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
	\label{fig:ps_q148_quadrant}
\end{figure}
